---
title: "Forensic Glass:  Exploratory Data Analysis"
output: html_document
---

efg, 2017-09-04

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
time.1 <- Sys.time()
```

Required packages

```{r, comment=NA, message=FALSE, warning=FALSE}
library(MASS)    # fgl
library(dplyr)   # select
library(caret)   # featurePlot
```

## Forensic Glass Data

See help(fgl) after loading MASS package.

RI = refractive index

Percentages by weight of oxides:
Na, Mg, Al, Si, K, Ca, Ba, Fe

type

* window float glass (WinF: 70), 
* window non-float glass (WinNF: 76), 
* vehicle window glass (Veh: 17), 
* containers (Con: 13), 
* tableware (Tabl: 9)
* vehicle headlamps (Head: 29).


```{r, comment=NA}
rawData <- fgl
str(rawData)
```

```{r, comment=NA}
head(rawData)
```


```{r, comment=NA}
table(rawData$type)
```

```{r, comment=NA}
type    <- rawData$type
rawData <- rawData %>% select(-type)
```

## First Look at Raw Data

```{r, comment=NA, fig.width=9, fig.height=9}
library(RColorBrewer)
colorSet <-brewer.pal(6, "Set2")   # Qualitative palette

par(mfrow=c(3,3))
plot(RI ~ type, data=fgl, col=colorSet, main="RI:  Refractive Index")
plot(Al ~ type, data=fgl, col=colorSet, main="Al:  Aluminum")
plot(Na ~ type, data=fgl, col=colorSet, main="Na:  Sodium")
plot(Mg ~ type, data=fgl, col=colorSet, main="Mg:  Magnesium")
plot(Ba ~ type, data=fgl, col=colorSet, main="Ba:  Barium")
plot(Si ~ type, data=fgl, col=colorSet, main="Si:  Silicon")
plot(K ~  type, data=fgl, col=colorSet, main="K:  Potasium")
plot(Ca ~ type, data=fgl, col=colorSet, main="Ca:  Calcium")
plot(Fe ~ type, data=fgl, col=colorSet, main="Fe:  Iron")
```

## Standardize Data

```{r, comment=NA}
transformSetup <- preProcess(rawData, method=c("center", "scale"))
centeredScaled <- predict(transformSetup, rawData)
```

## Data Visualization with Caret

Wrapper for Lattice Plotting of Predictor Variables

See [caret Github Page:  Visualizations](http://topepo.github.io/caret/visualizations.html)

### Boxplots

See [Box plot](https://en.wikipedia.org/wiki/Box_plot) at Wikipedia.

```{r, fig.width=12, fig.height=12}
featurePlot(centeredScaled, type, "box", main="Boxplots")
```

### Strip Plots

```{r, fig.width=12, fig.height=12}
featurePlot(centeredScaled, type,"strip", jitter=TRUE, main="Strip Plots")
```

### Pairs Plots

```{r, fig.width=12, fig.height=12}
featurePlot(centeredScaled, type, "pairs", auto.key=list(columns=6), main="Pairs Plots")
```

### Ellipse Plots

[This page](https://rstudio-pubs-static.s3.amazonaws.com/4991_14f069b28daf4d3187f3b5ae40051473.html) shows better examples of what ellipse plots show.

```{r, fig.width=12, fig.height=12}
featurePlot(centeredScaled, type, "ellipse", auto.key=list(columns=6), main="Ellipse Plots")
```

### Density Plots

See [Density estimation]().

```{r, fig.width=12, fig.height=12}
featurePlot(centeredScaled, type, "density", 
            scales=list(x=list(relation="free"), 
                        y=list(relation="free")),
            auto.key=list(columns=6), 
            main="Density Plots")
```

```{r, comment=NA, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2, time.1, units="secs"))), "sec\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H:%M:%S")`        

## References

Data also available through [UCI Repository](https://archive.ics.uci.edu/ml/datasets/glass+identification).

Forensic Glass data analysis is discussed in book [Data Mining and Business Analytics with R](https://books.google.com/books?id=gLnuxUmk5eEC&pg=PT133&lpg=PT133&dq=mass+forensic+glass+dataset&source=bl&ots=bQZsVGW1Xp&sig=zcxIBdyNXbLEdqz3XJNiaipIEus&hl=en&sa=X&ved=0ahUKEwip16CilYPWAhUM5IMKHVqACaIQ6AEISDAD#v=onepage&q=mass%20forensic%20glass%20dataset&f=false).

