---
title: "Forensic Glass:  glmnet:  LASSO, Ridge, and Elastic Net"
output: html_document
---

efg, 2017-09-06

[LASSO, Ridge, and Elastic Net](https://web.stanford.edu/~hastie/glmnet/glmnet_alpha.html) using [R glmnet package](https://cran.r-project.org/web/packages/glmnet/glmnet.pdf) with [caret](http://topepo.github.io/caret/index.html).

With [SMOTE:  Synthetic Minority Over-sampling Technique](https://www.jair.org/media/953/live-953-2037-jair.pdf)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
time.1 <- Sys.time()
```

### Required packages

```{r, comment=NA, message=FALSE, warning=FALSE}
library(MASS)          # fgl data
library(dplyr)         # select
library(caret)         # createDataParition, nearZeroVar
library(doParallel)    # registerDoParallel  
library(glmnet)
library(DMwR)          # SMOTE
library(gdata)
```

### Forensic Glass Data

```{r, comment=NA}
rawData <- fgl
dim(rawData)
```

```{r, comment=NA}
table(rawData$type)
prop.table( table(rawData$type) )
```

### Caret info about glmnet model

```{r, comment=NA}
getModelInfo()$glmnet$type
```

```{r, comment=NA}
getModelInfo()$glmnet$parameters
```

```{r, comment=NA}
getModelInfo()$glmnet$library
```

```{r, comment=NA}
getModelInfo()$glmnet$grid
```

### Define train and test datasets

```{r, comment=NA}
set.seed(71)

trainSetIndices <- createDataPartition(rawData$type, p=0.70, list=FALSE)

trainSet <- rawData[ trainSetIndices, ]
testSet  <- rawData[-trainSetIndices, ]
```

Assume dependent variable is last column of trainSet and testSet here.

```{r, comment=NA}
nrow(trainSet)
nrow(testSet)
```


```{r, comment=NA}
table(trainSet$type) 
table(testSet$type)
```


```{r, comment=NA}
prop.table( table(trainSet$type) )
prop.table( table(testSet$type) )
```

### SMOTE training data
```{r, comment=NA}
dim(trainSet)
table(trainSet$type)
```

```{r, comment=NA}
set.seed(1971)
trainSet <- SMOTE(type ~ ., trainSet, k=2, perc.over=800, perc.under=800)
dim(trainSet)
table(trainSet$type)
```

### Setup parallel processing

```{r, comment=NA}
rCluster <- makePSOCKcluster(6)   # use 6 cores
registerDoParallel(rCluster)  
```

## glmnet (lasso and elastic-net regularization) with SMOTE

```{r, comment=NA}
set.seed(29)
CVfolds   <-  5  # 5-fold cross validation (not enough data for 10 fold here)
CVrepeats <- 10  # repeat 10 times

# Used createMultiFolds to study 
indexFolds <- createMultiFolds(trainSet$type, CVfolds, CVrepeats)  # for repeated CV

trainControlParms <- trainControl(method = "repeatedcv",  # repeated cross validation
                                  number  = CVfolds,    
                                  repeats = CVrepeats,  
                                  index   = indexFolds,
                                  sampling = "smote",     ### sampling during resampling
                                  classProbs = TRUE,      # Estimate class probabilities
                                  summaryFunction = defaultSummary)

fit <- train(type ~ ., data=trainSet,
             preProcess = c("center", "scale"),
             method = "glmnet",
             metric = "Kappa",
             trControl = trainControlParms)

stopCluster(rCluster)
```

```{r, comment=NA}
plot(fit$finalModel)
```


```{r, comment=NA}
summary(fit$finalModel)
```

```{r, comment=NA}
print(fit)
```

### Variable Importance

See ?varImp

```{r, comment=NA, fig.width=8, fig.height=6}
plot( varImp(fit), main="Variable Importance" )
```

### Results on Train Set (In Sample)

Overly optimistic results for generalization

```{r, comment=NA}
options(width=120)
InSample  <- predict(fit, newdata=trainSet)
InSampleConfusion <- confusionMatrix(trainSet$type, InSample)
print(InSampleConfusion)   
```

### Results on Test Set (Out of Sample)

More realistic results on predictions with new data

```{r, comment=NA}
options(width=120)
OutOfSample  <- predict(fit, newdata=testSet)
confusion <- confusionMatrix(testSet$type, OutOfSample)
print(confusion)   
```

```{r, comment=NA, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2, time.1, units="secs"))), "sec\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H:%M:%S")`   

## References

* [Glmnet Vignette](https://web.stanford.edu/~hastie/glmnet/glmnet_alpha.html), Trevor Hastie and Junyang Qian, 2014.

* [Trevor Hastie presents glmnet: lasso and elastic-net regularization in R](https://www.r-bloggers.com/trevor-hastie-presents-glmnet-lasso-and-elastic-net-regularization-in-r/),  R Bloggers, 2013.

* [An Introduction to Statistical Learning with Applications in R](http://www-bcf.usc.edu/~gareth/ISL/), Gareth James, Daniela Witten, Trevor Hastie, Robert Tibshirani, 2013
