---
title: "Forensic Glass:  Singular Value Decomposition"
output: html_document
---

efg, 2017-09-02

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Setup

```{r}
time.1 <- Sys.time()
```

Required packages

```{r, comment=NA, message=FALSE, warning=FALSE}
library(MASS)      # fgl data
library(dplyr)     # select
```

Forensic Glass Data

```{r, comment=NA}
rawData <- fgl
rawData  <- rawData %>% select(-type) %>% as.matrix()

# Mean=0, StdDev=1
centeredScaledData <- scale(rawData, center=TRUE, scale=TRUE)
```

## Singular Value Decomposition (SVD) 

Decompose original matrix into product: u * diag(d) * t(v)

Singular values, also known as the eigenvalues, are along diag(d) in decreasing order.

```{r, comment=NA}
udv <- svd(centeredScaledData)   # centered, scaled data
str(udv)
```

### Variance explained by EigenFeatures

The total variance is the sum of the squares of the eigenvalues -- the diagnonal terms of the diag(d) scaling matrix.

```{r, comment=NA,fig.width=7, fig.height=7}
plot(cumsum(udv$d^2 / sum(udv$d^2)), type="b", pch=15, ylim=c(0,1), col="blue",
     main="Scree Plot", xlab="Principal Component", ylab="Fraction Variance Explained",
     xaxt="n")
axis(1, 1:length(udv$d), paste0("PC", 1:length(udv$d)) )
grid()

lines(udv$d^2 / sum(udv$d^2), type="b", pch=20, ylim=c(0,1), col="black")
legend("right", c("Cumulative Variance", "Variance"), pch=c(15, 20), col=c("blue", "black"), box.lty=0)
```

Variance explained by eigenvalue

```{r, comment=NA}
udv$d^2 / sum(udv$d^2)  %>% round(4)
```

Cumlative variance explained

```{r, comment=NA}
cumsum(udv$d^2 / sum(udv$d^2))  %>% round(4)
```

So, the first three singular values explain about 66% of the variance -- the first four explain almost 80%.

### Right Singular Vectors Overview

```{r, comment=NA}
boxplot(udv$v, main="Right singular vectors", col="skyblue")
grid()
```

Variable weighting by right singular vector

### 1st right singular vector

Moderate high Al with low RI and Ca

```{r, comment=NA}
N <- nrow(udv$v)
x <- barplot(udv$v[,1], main="Right singular vector 1",
             width=rep(0.8,N), space=rep(0.2, N))
text(x, udv$v[,1]/2, colnames(rawData), adj=0.5)
abline(h=0)
grid()
```

### 2nd  right singular vector

Low Mg with moderate high Ba

```{r, comment=NA}
x <- barplot(udv$v[,2], main="Right singular vector 2",
             width=rep(0.8,N), space=rep(0.2, N))
text(x, udv$v[,2]/2, colnames(rawData), adj=0.5)
abline(h=0)
grid()
```

### 3rd  right singular vector

Low K with moderate high SI

```{r, comment=NA}
x <- barplot(udv$v[,3], main="Right singular vector 3",
             width=rep(0.8,N), space=rep(0.2, N))
text(x, udv$v[,3]/2, colnames(rawData), adj=0.5)
abline(h=0)
grid()
```

### 4th  right singular vector

High Si with moderate low Na and Mg
 
```{r, comment=NA}
x <- barplot(udv$v[,4], main="Right singular vector 4",
             width=rep(0.8,N), space=rep(0.2, N))
text(x, udv$v[,4]/2, colnames(rawData), adj=0.5)
abline(h=0)
grid()
```   

### SVD Technical Notes

The singular value decomposition of matrix M is as follows:

   
$M = U * \Sigma * V'$   

where 

$U$ = left singular vectors

$\Sigma$ = singular values diagonal matrix. The diagonal terms are the singular values, usually listed in decreasing order. The singular values are the eigenvalues of the matrix. 

$V$ = right singular vectors 

In R, given

```
    udv <- svd(M)
```

The original matrix M can be recomputed as follows: 

```
    M <- udv$u %*% diag(udv$d) %*% t(udv$v)
```
    
where t() is the transpose function.

#### Fewer dimensions

Much of the variance of matrix M can be captured using fewer features (dimensions) based on the variance explained by the singular values.


$M_{approx} = U_n * \Sigma_n * V_n'$   

where

$U_n$ = first n columns of U left singular vectors, udv$u[,1:features] in R

$\Sigma_n$ = first n diagonal terms of $\Sigma$ matrix, diag(udv$d[1:features]) in R. 

$V_n$ = first n columns of V right singular vectors, udv$v[,1:features] in R. 

For the first four eigenvalues that explain about 80% of variance:

```{r, comment=NA}
    features <- 4
    Mapprox <- udv$u[,1:features] %*% diag(udv$d[1:features]) %*% t(udv$v[,1:features])
```

Maximum absolute error between approximation and original value

```{r, comment=NA}
max(abs(centeredScaledData - Mapprox))
```

For features = 9, the original matrix and approximation match within machine precision.

#### PCA scores

PCA scores can be computed from the SVD U vectors and the diagnoal scaling matrix:

```{r, comment=NA}
PCAscores <- udv$u %*% diag(udv$d)  # Principal components
dim(PCAscores)
```

Let's display the first 3 principal components, which will be used in the PCA example.

```{r, comment=NA}
PCAscores[, 1:3]  %>% head
```

Compare these values to those created using the caret's preProcess function in the separate PCA notebook.

```{r, comment=NA, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2, time.1, units="secs"))), "sec\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H:%M:%S")`      

## References

[Singular Value Decomposition](https://en.wikipedia.org/wiki/Singular_value_decomposition), Wikipedia