---
title: "Forensic Glass:  Boruta 'All Relevant' Variables"
output: html_document
---

efg, 2017-09-04

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
time.1 <- Sys.time()
```

Required packages

```{r, comment=NA, message=FALSE, warning=FALSE}
library(MASS)          # fgl data
library(dplyr)         # select
library(doParallel)    # registerDoParallel  
library(Boruta)        # Bortua
library(tibble)        # rownames_to_column  
```

Forensic Glass Data

```{r, comment=NA}
rawData <- fgl
```

Setup parallel processing

```{r, comment=NA}
rCluster <- makePSOCKcluster(6)   # use 6 cores
registerDoParallel(rCluster)  
```

## Boruta 'All Relevant' Variables

```{r, comment=NA}
set.seed(71)
BorutaModel <- Boruta(type ~ ., data=rawData,
                      getImp=getImpFerns,
                      maxRuns=1000, num.threads=6)
stopCluster(rCluster)
```

```{r, comment=NA}
print(BorutaModel)
```

```{r, comment=NA, fig.width=10, fig.height=8}
stats <- attStats(BorutaModel)
statsOrdered <- stats[order(stats$medianImp, decreasing=TRUE), ]

plot(BorutaModel, las=2, cex.axis=0.75, main="Boruta Importance")
```

```{r, comment=NA, fig.width=10, fig.height=8}
plotImpHistory(BorutaModel, main="Importance History")
```

```{r, comment=NA}
BorutaFeatures <- statsOrdered %>% rownames_to_column(var="Feature")
print(head(BorutaFeatures))                                                         
```


```{r, comment=NA, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2, time.1, units="secs"))), "sec\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H:%M:%S")`        
    
## References

* [Feature Selection with the Boruta Package](https://www.jstatsoft.org/article/view/v036i11/v36i11.pdf), Miron B. Kursa and Witold R. Rudnicki,  *Journal of Statistical Software*, 2010.

* [Feature selection: All-relevant selection with the Boruta package](https://www.r-bloggers.com/feature-selection-all-relevant-selection-with-the-boruta-package/), Allan Engelhardt, 2010.

* Venn Diagram Comparison of Boruta, FSelectorRcpp and GLMnet Algorithms(http://r-addict.com/2016/06/19/Venn-Diagram-RTCGA-Feature-Selection.html), JMarcin Kosiński, 2016.