---
title: "Forensic Glass:  Heatmap and Hierarchical Clustering"
output: html_document
---

efg, 2017-09-02

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r, comment=NA}
time.1 <- Sys.time()
```

Required packages

```{r, comment=NA, message=FALSE, warning=FALSE}
library(MASS)         # fgl data
library(NMF)          # aheatmap
library(dplyr)        # select
library(doParallel)   # parallel processing:  registerDoParallel
library(pvclust)      # pvclust
```

Forensic Glass Data

```{r, comment=NA}
rawData <- fgl
rownames(rawData) <- make.names( as.character(rawData$type), unique=TRUE )  # unique rownames
rowLabel <- rawData["type"]   # data.frame with single column "type"
rawData  <- rawData %>% select(-type) %>% as.matrix()

# Mean=0, StdDev=1
centeredScaledData <- scale(rawData, center=TRUE, scale=TRUE)
```

## Heatmap 
```{r, comment=NA, fig.width=8, fig.height=16}
colorScale <- colorRampPalette(c("#7F0000","red","#FF7F00","yellow","white",
                                 "cyan", "#007FFF", "blue","#00007F"))(100)  

aheatmap(centeredScaledData, color=colorScale,       # Diverging Palette
         Rowv=NA, Colv=NA,                           # no re-ordering of data
         scale="col", breaks=0,                      # Palette centered on this value
         annRow=rowLabel, annColors="Dark2",         # Qualitative Palette (RColorBrewer)
         treeheight=c(100, 100),                     # size of dendrograms
         fontsize=14, cexCol=1.25, 
         width=8, height=16)
```

## Clustering of Predictors and Instances

```{r, comment=NA, fig.width=8, fig.height=16}
aheatmap(centeredScaledData, color=colorScale,       # Diverging Palette
         scale="col", breaks=0,                      # Palette centered on this value
         annRow=rowLabel, annColors="Dark2",         # Qualitative Palette (RColorBrewer)
         distfun="pearson",                          # start wtih "pearson" or "euclidean"
         treeheight=c(100, 100),                     # size of dendrograms
         fontsize=14, cexCol=1.25, 
         width=8, height=16)
```

This is a visual view of the problem.  The vertical color bar at the left shows the glass type for the predictors to the right.  What patterns of predictors can be observed for a given glass type?

There are many eqivalent ways to display the [dendrogram](https://en.wikipedia.org/wiki/Dendrogram) trees above.  Any pair of nodes shown connected in the dendrogram can be flipped for an equivalent representation.

The "Head" type formed a tight cluster at the top, but other types do not appear in such tight clusters.

How reliable are the clusters of predictors shown in the dendrogram at the top? 

### Evaluate Uncertainty in Clustering Predictors

There are many equivalent trees.  How reliable are the branches of predictors if we look at many clusterings?

**pvclust** Algorithm

1. Generated thousands of bootstrap samples by randomly sampling elements of the data

1. Compute hierarchical clustering on each bootstrap copy

1. For each cluster:

* compute the **bootstrap probability (BP)** value which corresponds to the frequency that the cluster is identified in bootstrap copies.

* Compute the **approximately unbiased (AU)** probability values (p-values) by multiscale bootstrap resampling.

```{r, comment=NA, message=FALSE, warning=FALSE}
rCluster <- makePSOCKcluster(6)   # use 6 cores
registerDoParallel(rCluster)
```

Use nboot=1000 for testing at first, and then use nboot=10000 for smaller error.

```{r, comment=NA, message=FALSE, warning=FALSE}
pvResult <- pvclust(centeredScaledData,
                    parallel=TRUE,
                    method.hclust="ward.D2", method.dist="correlation",
                    nboot=10000, iseed=19937)
stopCluster(rCluster)
```

```{r, comment=NA, fig.width=10, fig.height=8}
plot(pvResult, cex=0.8, cex.pv=0.7)
pvrect(pvResult, alpha=0.95)
```

**Clusters with AU > = 95% are considered to be strongly supported by data.**

*For a cluster with AU p-value > 0.95, the hypothesis that "the cluster does not exist" is rejected with significance level 0.05.*

```{r, comment=NA}
pvResult
```

```{r, comment=NA, message=FALSE, warning=FALSE}
seplot(pvResult)
```

Investigate any clusters with extremely large standard errors.

```{r, comment=NA, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2, time.1, units="secs"))), "sec\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H:%M:%S")`         

## References

* [NMF: Algorithms and Framework for Nonnegative Matrix Factorization (NMF)](https://cran.r-project.org/web/packages/NMF/index.html), CRAN.

* [Generating heatmaps for Nonnegative Matrix Factorization](https://cran.r-project.org/web/packages/NMF/vignettes/heatmaps.pdf), CRAN.

* [Pvclust: an R package for assessing the uncertainty in hierarchical clustering](https://www.ncbi.nlm.nih.gov/pubmed/16595560), *Bioinformatics*, 2006.

* [pvclust:  An R package for hierarchical clustering with p-values](http://stat.sys.i.kyoto-u.ac.jp/prog/pvclust/), Ryota Suzuki and Hidetoshi Shimodaira.
