---
title: "Forensic Glass:  C5.0 Classification"
output: html_document
---

efg, 2017-09-04

[C5.0 classification](https://www.rulequest.com/see5-unix.html) using [C50 package](https://cran.r-project.org/web/packages/C50/C50.pdf) with [caret](http://topepo.github.io/caret/index.html).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
time.1 <- Sys.time()
```

### Required packages

```{r, comment=NA, message=FALSE, warning=FALSE}
library(MASS)          # fgl data
library(dplyr)         # select
library(caret)         # createDataParition, nearZeroVar
library(doParallel)    # registerDoParallel  
library(C50)
```

### Forensic Glass Data

```{r, comment=NA}
rawData <- fgl
dim(rawData)
```

```{r, comment=NA}
table(rawData$type)
prop.table( table(rawData$type) )
```

### Caret info about C5.0 model

```{r, comment=NA}
getModelInfo()$C5.0$type
```

```{r, comment=NA}
getModelInfo()$C5.0$parameters
```


```{r, comment=NA}
getModelInfo()$C5.0$library
```

```{r, comment=NA}
getModelInfo()$C5.0$grid
```

### Define train and test datasets

```{r, comment=NA}
set.seed(71)

trainSetIndices <- createDataPartition(rawData$type, p=0.70, list=FALSE)

trainSet <- rawData[ trainSetIndices, ]
testSet  <- rawData[-trainSetIndices, ]
```

Assume dependent variable is last column of trainSet and testSet here.

```{r, comment=NA}
nrow(trainSet)
nrow(testSet)
```

```{r, comment=NA}
prop.table( table(trainSet$type) )
prop.table( table(testSet$type) )
```
### Setup parallel processing

```{r, comment=NA}
rCluster <- makePSOCKcluster(6)   # use 6 cores
registerDoParallel(rCluster)  
```

## C5.0

```{r, comment=NA, message=FALSE, warning=FALSE}
set.seed(29)
CVfolds   <-  5  # 5-fold cross validation (not enough data for 10 fold here)
CVrepeats <- 10  # repeat 10 times

trainControlParms <- trainControl(method = "repeatedcv",  # repeated cross validation
                                  number  = CVfolds,    
                                  repeats = CVrepeats,  
                                  classProbs = TRUE,      # Estimate class probabilities
                                  summaryFunction = defaultSummary)

fit <- train(type ~ ., data=trainSet,
             method = "C5.0", 
             tuneLength = 11,
             metric = "Kappa",                 # helps with imbalance
             trControl = trainControlParms)

stopCluster(rCluster)
```

```{r, comment=NA}
print(fit$finalModel)
```

```{r, comment=NA}
summary(fit$finalModel)
```

```{r, comment=NA}
print(fit)
```

```{r, comment=NA}
plot(fit)
```

### Variable Importance

See ?varImp

```{r, comment=NA, fig.width=8, fig.height=6}
plot( varImp(fit), main="Variable Importance" )
```

### Results on Train Set (In Sample)

Overly optimistic results for generalization

```{r, comment=NA}
options(width=120)
InSample  <- predict(fit, newdata=trainSet)
InSampleConfusion <- confusionMatrix(trainSet$type, InSample)
print(InSampleConfusion)   
```

### Results on Test Set (Out of Sample)

More realistic results on predictions with new data

```{r, comment=NA}
options(width=120)
OutOfSample  <- predict(fit, newdata=testSet)
confusion <- confusionMatrix(testSet$type, OutOfSample)
print(confusion)   
```

```{r, comment=NA, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2, time.1, units="secs"))), "sec\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H:%M:%S")`   

## References

* [Classification Using C5.0](https://static1.squarespace.com/static/51156277e4b0b8b2ffe11c00/t/51e7e42ce4b0fd2e32684bca/1374151724529/user_C5.0.pdf), Max Kuhn, UseR! 2013.

* [Understanding the output of C5.0 classification model using the CARET package](https://stats.stackexchange.com/questions/83913/understanding-the-output-of-c5-0-classification-model-using-the-caret-package), *Cross Validated*, 2014.

* [R, caret, and Parameter Tuning C5.0 ](http://www.euclidean.com/machine-learning-in-practice/2015/6/12/r-caret-and-parameter-tuning-c50), John Alberg, 2015.

