---
title: "Forensic Glass:  Correlation Matrix"
output: html_document
---

efg, 2017-09-02

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
time.1 <- Sys.time()
```

Required packages

```{r, comment=NA, message=FALSE, warning=FALSE}
library(MASS)      # fgl data
library(corrplot)  # corrplot
library(caret)     # findCorrelation
library(dplyr)     # select
```

Forensic Glass Data

```{r, comment=NA}
rawData <- fgl %>% select(-type)
head(rawData)
```

## Correlation Matrix

```{r, fig.width=8, fig.height=8}
colorScale <- colorRampPalette(c("#7F0000","red","#FF7F00","yellow","white",
                                 "cyan", "#007FFF", "blue","#00007F"))(100)  

corMatrix <- cor(rawData)  # Create correlation matrix

corrplot(corMatrix, type="upper", method="square", order="AOE",
         tl.col="black", tl.srt=45, tl.cex=1.5,
         addCoef.col="darkgrey", number.cex=1.25,
         col=colorScale)
mtext("Correlation Matrix (angular order of eigenvectors)", line=3)         
```

## Removing high correlation pairs

Some machine learning algorithms are impeded by highly correlated predictors.  Caret's *findCorrelation* procedure can be used to remove one of the highly-correlated pairs.

```{r, comment=NA}
HIGH_CORRELATION_CUTOFF <- 0.80

corHigh <- findCorrelation(corMatrix, HIGH_CORRELATION_CUTOFF)
if (length(corHigh) > 0)
{
  cat("Removing highly-correlated variable(s): ", names(rawData)[corHigh])
  rawData <- rawData[, -corHigh]  
}
head(rawData)
```

In most machine learning applications, the *findCorrelation* check would be applied to the training set data, but any highly-correlated predictors would be removed from both the training and test sets.

```{r, comment=NA, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2, time.1, units="secs"))), "sec\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H:%M:%S")`         

## References

* [corrplot: Visualization of a Correlation Matrix](https://cran.r-project.org/web/packages/corrplot/index.html), CRAN.
    
* [An Introduction to corrplot Package](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html), CRAN.

* [Visualize correlation matrix using correlogram](http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram), STHDA.

* [An Introduction to Matrix Visualization & corrplot Package](https://uploads.cosx.org/2009/12/An-Introduction-to-Matrix-Visualization-and-corrplot-Package.pdf), Taiyun Wei, 2009.

* [Introduction to Feature selection for bioinformaticians using R, correlation matrix filters, PCA & backward selection](Introduction to Feature selection for bioinformaticians using R, correlation matrix filters, PCA & backward selection), Yasset Perez-Riverol, 2013.