---
title: "Forensic Glass:  Neural Network Classification (nnet)"
output: html_document
---

efg, 2017-09-04

[Neural Networks](https://visualstudiomagazine.com/articles/2016/11/01/using-the-r-nnet-package.aspx) using [nnet package](https://cran.r-project.org/web/packages/nnet/nnet.pdf) with [caret](http://topepo.github.io/caret/index.html).

nnet provides feed-forward neural networks with a single hidden layer.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
time.1 <- Sys.time()
```

### Required packages

```{r, comment=NA, message=FALSE, warning=FALSE}
library(MASS)          # fgl data
library(dplyr)         # select
library(caret)         # createDataParition, nearZeroVar
library(doParallel)    # registerDoParallel  
library(nnet)
```

Special setup to plot neural network.  Import the plot.nnet function from Github
```{r, comment=NA}
library(devtools)
source_url('https://gist.githubusercontent.com/fawda123/7471137/raw/466c1474d0a505ff044412703516c34f1a4684a5/nnet_plot_update.r')
```

### Forensic Glass Data

```{r, comment=NA}
rawData <- fgl
dim(rawData)
```

```{r, comment=NA}
table(rawData$type)
prop.table( table(rawData$type) )
```

### Caret info about neural network (nnet) model

```{r, comment=NA}
getModelInfo()$nnet$type
```

```{r, comment=NA}
getModelInfo()$nnet$parameters
```

```{r, comment=NA}
getModelInfo()$nnet$library
```

```{r, comment=NA}
getModelInfo()$nnet$grid
```

### Define train and test datasets

```{r, comment=NA}
set.seed(71)

trainSetIndices <- createDataPartition(rawData$type, p=0.70, list=FALSE)

trainSet <- rawData[ trainSetIndices, ]
testSet  <- rawData[-trainSetIndices, ]
```

Assume dependent variable is last column of trainSet and testSet here.

```{r, comment=NA}
nrow(trainSet)
nrow(testSet)
```


```{r, comment=NA}
table(trainSet$type) 
table(testSet$type)
```


```{r, comment=NA}
prop.table( table(trainSet$type) )
prop.table( table(testSet$type) )
```

### Setup parallel processing

```{r, comment=NA}
rCluster <- makePSOCKcluster(6)   # use 6 cores
registerDoParallel(rCluster)  
```

## Neural Network

```{r, comment=NA}
set.seed(29)
CVfolds   <-  5  # 5-fold cross validation (not enough data for 10 fold here)
CVrepeats <- 10  # repeat 10 times

trainControlParms <- trainControl(method = "repeatedcv",  # repeated cross validation
                                  number  = CVfolds,    
                                  repeats = CVrepeats,
                                  classProbs = TRUE,        # Estimate class probabilities
                                  summaryFunction = defaultSummary)

fit <- train(type ~ ., data=trainSet,
             preProcess = c("center", "scale"),
             method = "nnet",
             tuneLength = 15,                    #####
             metric = "Kappa",                   #####
             linout = TRUE,                      #####
             trControl=trainControlParms)

stopCluster(rCluster)
```

```{r, comment=NA}
print(fit$finalModel)
```

```{r, comment=NA, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
plot.nnet(fit$finalModel)
```

```{r, comment=NA}
summary(fit$finalModel)
```

```{r, comment=NA}
print(fit)
```

```{r, comment=NA}
plot(fit)
```


### Variable Importance

Unclear why cannot plot varImp

```
plot( varImp(fit), main="Variable Importance" )
```

```{r, comment=NA}
print( varImp(fit), main="Variable Importance" )
```

### Results on Train Set (In Sample)

Overly optimistic results for generalization

```{r, comment=NA}
options(width=120)
InSample  <- predict(fit, newdata=trainSet)
InSampleConfusion <- confusionMatrix(trainSet$type, InSample)
print(InSampleConfusion)   
```

### Results on Test Set (Out of Sample)

More realistic results on predictions with new data

```{r, comment=NA}
options(width=120)
OutOfSample  <- predict(fit, newdata=testSet)
confusion <- confusionMatrix(testSet$type, OutOfSample)
print(confusion)   
```

```{r, comment=NA, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2, time.1, units="secs"))), "sec\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H:%M:%S")`   

## References

* [How to model a neural network through the use of caret, R](https://stackoverflow.com/questions/21622975/how-to-model-a-neural-network-through-the-use-of-caret-r), Stack Overflow, 2014.

* [Visualizing neural networks in R – update](https://beckmw.wordpress.com/tag/nnet/), R is my friend blog
