---
title: "Forensic Glass:  Linear Discriminant Analysis"
output: html_document
---

efg, 2017-09-04

[Linear Discriminant Analysis](https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/lda.html) using [R MASS package](https://cran.r-project.org/web/packages/MASS/MASS.pdf) with [caret](http://topepo.github.io/caret/index.html).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
time.1 <- Sys.time()
```

### Required packages

```{r, comment=NA, message=FALSE, warning=FALSE}
library(MASS)          # fgl data, lda
library(dplyr)         # select
library(caret)         # createDataParition, nearZeroVar
library(doParallel)    # registerDoParallel  
```

### Forensic Glass Data

```{r, comment=NA}
rawData <- fgl
dim(rawData)
```

```{r, comment=NA}
table(rawData$type)
prop.table( table(rawData$type) )
```

### Caret info about linear discriminant analysis (lda) model

```{r, comment=NA}
getModelInfo()$lda$type
```

```{r, comment=NA}
getModelInfo()$lda$parameters
```

```{r, comment=NA}
getModelInfo()$lda$library
```

```{r, comment=NA}
getModelInfo()$lda$grid
```

### Define train and test datasets

```{r, comment=NA}
set.seed(71)

trainSetIndices <- createDataPartition(rawData$type, p=0.70, list=FALSE)

trainSet <- rawData[ trainSetIndices, ]
testSet  <- rawData[-trainSetIndices, ]
```

Assume dependent variable is last column of trainSet and testSet here.

```{r, comment=NA}
nrow(trainSet)
nrow(testSet)
```


```{r, comment=NA}
table(trainSet$type) 
table(testSet$type)
```


```{r, comment=NA}
prop.table( table(trainSet$type) )
prop.table( table(testSet$type) )
```

#### Remove near-zero variance predictors

```{r, comment=NA}
nearZero <- nearZeroVar(trainSet, saveMetrics=TRUE)
nearZero
```
Eliminate variables that have zero variance (zeroVar), or variables with near zero variance (nzv).  

In this case, we'll remove any near zero variance variables.

```{r, comment=NA}
nzvCount <- sum(nearZero$nzv)
if (nzvCount > 0)
{
  trainSet <- trainSet[, !nearZero$nzv]  
  testSet  <- testSet[,  !nearZero$nzv]  # make same change to testSet
}
```

#### Remove variables with high correlation to others

```{r, comment=NA}
HIGH_CORRELATION_CUTOFF <- 0.90
corMatrix <- trainSet %>% select(-type) %>% cor   # create correlation matrix

corHigh <- findCorrelation(corMatrix, HIGH_CORRELATION_CUTOFF)
if (length(corHigh) > 0)
{
  cat("Removing highly-correlated variable(s): ", names(trainSet)[corHigh])
  trainSet <- trainSet[, -corHigh]  
  testSet  <- testSet[,  -corHigh]  
}
```

#### Remove variables with linear dependencies

QR decomposition is used to determine if the matrix is full rank.  Sets of columns that are involved in the dependencies are identified.

```{r, comment=NA}
linearCombos <- findLinearCombos(trainSet %>% select(-type))
if (length(linearCombos$remove) > 0)
{
  cat("Removing linear dependencies: ", names(trainSet)[linearCombos$remove])
  trainSet <- trainSet[, -linearCombos$remove]  
  testSet  <- testSet[,  -linearCombos$remove]  
}
```

### Setup parallel processing

```{r, comment=NA}
rCluster <- makePSOCKcluster(6)   # use 6 cores
registerDoParallel(rCluster)  
```

## Linear Discriminant Analysis 

```{r, comment=NA}
set.seed(29)
CVfolds   <-  5  # 5-fold cross validation (not enough data for 10 fold here)
CVrepeats <- 10  # repeat 10 times

# Used createMultiFolds to study 
indexFolds <- createMultiFolds(trainSet$type, CVfolds, CVrepeats)  # for repeated CV

trainControlParms <- trainControl(method = "repeatedcv",  # repeated cross validation
                                  number  = CVfolds,    
                                  repeats = CVrepeats,  
                                  index   = indexFolds,
                                  classProbs = TRUE,        # Estimate class probabilities
                                  summaryFunction = defaultSummary)

fit <- train(type ~ ., data=trainSet,
             preProcess = c("center", "scale"),
             method = "lda",
             metric = "Kappa",
             trControl = trainControlParms)

stopCluster(rCluster)
```

```{r, comment=NA}
print(fit$finalModel)
```

```{r, comment=NA}
print(fit)
```

### Variable Importance

See ?varImp

```{r, comment=NA, fig.width=8, fig.height=6}
plot( varImp(fit), main="Variable Importance" )
```

### Results on Train Set (In Sample)

Overly optimistic results for generalization

```{r, comment=NA}
options(width=120)
InSample  <- predict(fit, newdata=trainSet)
InSampleConfusion <- confusionMatrix(trainSet$type, InSample)
print(InSampleConfusion)   
```

### Results on Test Set (Out of Sample)

More realistic results on predictions with new data

```{r, comment=NA}
options(width=120)
OutOfSample  <- predict(fit, newdata=testSet)
confusion <- confusionMatrix(testSet$type, OutOfSample)
print(confusion)   
```

```{r, comment=NA, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2, time.1, units="secs"))), "sec\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H:%M:%S")`   

## References

* [Feature Selection with the Caret R Package](https://machinelearningmastery.com/feature-selection-with-the-caret-r-package/), Jason Brownlee, 2014.

* [How to Evaluate Machine Learning Algorithms with R](https://machinelearningmastery.com/evaluate-machine-learning-algorithms-with-r/), Jason Brownlee, 2016.  See section *Algorithm Spot Check* for getting started with a variety of algorithms.

* [Get Your Data Ready For Machine Learning in R with Pre-Processing](https://machinelearningmastery.com/pre-process-your-dataset-in-r/), Jason Brownlee, 2016.

